---
title: "Open Street Map"
subtitle: "Egypt, May-2021"
author: "Hany Nagaty"
output:
  pdf_document:
     latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, message = FALSE)
```

```{r}
library(sf)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
```
# Open Street Maps Description
This is a brief description for the data downloaded from [Open Street Map](https://www.openstreetmap.org/#map=13/30.0225/31.4812) website. OSM is described as “a map of the world, created by people like you and free to use under an open license”.
I downloaded the data for Egypt on May-2021. I used this [mirror](https://download.geofabrik.de/africa.html).


```{r shpTable, eval=FALSE}
shp %>%
  as_tibble() %>% 
  mutate(fclass = as.factor(fclass)) %>% 
  mutate(fclass = fct_lump_n(fclass, 5)) %>% 
  pull(fclass) %>% 
  fct_count(prop = TRUE) %>% 
  mutate(p = percent(p)) %>% 
  rename(Class = f, Count = n, Proportion = p) %>% 
  arrange(desc(Count)) %>% 
  kable(digits = 2, 
        format.args = list(big.mark = ","),
        caption = "Counts Table") %>% 
  kable_classic(full_width = FALSE,  position = "left") %>% 
  column_spec(3, bold = TRUE)
```

```{r shpPlot, eval=FALSE}
ggplot(shp) +
  geom_sf(aes(color = fclass)) +
  theme(legend.position = "none")
```

### The data has the below **16 layers**    

## 1. `gis_osm_roads_free_1`

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_roads_free_1.shp"
shp <- st_read(shp_file)
```

This is the **road network**. It has classes such as motorway,  primary, secondary, residential, …..etc.  

Total number of line segments: `r nrow(shp)`.  

```{r, include=TRUE, ref.label=c('shpTable')}
```

``` {r, include=TRUE}
shp %>% 
  as_tibble() %>%
  count(bridge, sort = TRUE) %>%
  kable(digits = 2, format.args = list(big.mark = ",")) %>% 
  kable_classic(full_width = FALSE,  position = "left") %>% 
  column_spec(2, bold = TRUE)
shp %>% 
  as_tibble() %>%
  count(tunnel, sort = TRUE) %>%
  kable(digits = 2, format.args = list(big.mark = ",")) %>% 
  kable_classic(full_width = FALSE,  position = "left") %>% 
  column_spec(2, bold = TRUE)
shp %>% 
  as_tibble() %>%
  count(oneway, sort = TRUE) %>%
  kable(digits = 2, format.args = list(big.mark = ",")) %>% 
  kable_classic(full_width = FALSE,  position = "left") %>% 
  column_spec(2, bold = TRUE)
```

```{r, include=TRUE, ref.label='shpPlot'}
```



## 2. `gis_osm_pois_free_1`

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_pois_free_1.shp"
shp <- st_read(shp_file)
```
**Points of Interest**. This might include shopping places, hotels, hospitals, water towers, telecom towers.   
*Note that the list is not comprehensive*  

Total number of points: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpTable', 'shpPlot')}
```


## 3. `gis_osm_pois_a_free_1`

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_pois_a_free_1.shp"
shp <- st_read(shp_file)
```
Another set of **Points of Interest**.  
In contrast to the previous list, this list contains MULTIPOLYGON instead of POINT.  

Total number of points: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```


## 4. gis_osm_pofw_free_1

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_pofw_free_1.shp"
shp <- st_read(shp_file)
```
**Places of Worship**, this includes mosques and churches.  

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```


## 5. gis_osm_pofw_a_free_1

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_pofw_a_free_1.shp"
shp <- st_read(shp_file)
```
Another set of **Places of Worship**. The GIS features here are in POLYGON format rather than POINT format.    

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```


## 6. gis_osm_railways_free_1

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_railways_free_1.shp"
shp <- st_read(shp_file)
```

This is the **railway network**. It has classes such as railway, tram and subway.  

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```


## 7. gis_osm_traffic_free_1

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_traffic_free_1.shp"
shp <- st_read(shp_file)
```

This is the **traffic landmarks**, such as crossings or junctions.  

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```


## 8. gis_osm_traffic_a_free_1

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_traffic_a_free_1.shp"
shp <- st_read(shp_file)
```

Another set of **traffic landmarks**. The GIS features here are in MULTIPOLYGON format rather than POINT format.  

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```


## 9. gis_osm_transport_free_1

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_transport_free_1.shp"
shp <- st_read(shp_file)
```

**Transportation landmarks**, this includes bus stations and others.   
*The list is not comprehensive*  

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```


## 10. gis_osm_transport_a_free_1

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_transport_a_free_1.shp"
shp <- st_read(shp_file)
```

Another set of **transportation landmarks**. The GIS features here are in MULTIPOLYGON format rather than POINT format.  
*The list is not comprehensive*  

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```

  
## 11. gis_osm_buildings_a_free_1

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_buildings_a_free_1.shp"
shp <- st_read(shp_file)
```

Building boundaries. Some buildings has names; I found places such as تلفزيون وراديو ماسبيرو, مستشفى العجوزة.   
*This layer is very sparse and lacks most of the buildings.*  

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```

  
## 12. gis_osm_places_free_1

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_places_free_1.shp"
shp <- st_read(shp_file)
```

**Places**. It contains villages and others.  

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```
  

## 13. `gis_osm_places_a_free_1`

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_places_a_free_1.shp"
shp <- st_read(shp_file)
```

**Places** in POLYGON rather than POINT format.  

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```

  
## 14. `gis_osm_natural_free_1`

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_natural_free_1.shp"
shp <- st_read(shp_file)
```

**Natural** landmarks, like trees or springs.  

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```


## 15. `gis_osm_natural_a_free_1`

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_natural_a_free_1.shp"
shp <- st_read(shp_file)
```

**Natural** landmarks in POLYGON rather than POINT format.    

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```


## 16. `gis_osm_water_a_free_1`

```{r}
shp_file <- "D:/Geodata/~MapsInUse/EgyptOSM/gis_osm_water_a_free_1.shp"
shp <- st_read(shp_file)
```

**Water** features, it includes the River Nile, lakes, ... etc.    

Total number of features: `r nrow(shp)`

```{r, include=TRUE, ref.label=c('shpRead', 'shpTable', 'shpPlot')}
```

  
